Astr 513: Homework 1

### 2016-02-11 HW1/submission/CMW_hw1_v2.tex
Added problem 2 to the tex file.
I can’t compile this on my own to make sure it compiles correctly,
because I’m missing some of the packages apparently. If it doesn’t
compile properly, let me know.

\section{Designing surveys}

>>>>

\section{Designing surveys}
We are given a prior probability from the rate of detecting planets of $1-2 R_{\oplus}$ around solar type stars from Kepler, and are told to assume this rate has been measured to high accuracy as being 10\%. We are tasked with determining the number of solar type stars we would need to observe to find at least 30 $1-2 R_{\oplus}$ planets in a similar survey observing a different area of the sky with a 90\% and 99\% likelihood. To do this, we will want to make of the binomial distribution,
\begin{equation*}
  P(n, \rho | N) = \left(N \choose n\right) \rho^n \left(1 - \rho\right)^{N - n},
\end{equation*}
where $n$ is the number of successes, $\rho$ is the prior likelihood of obtaining a success, and $N$ is the number of independent observations. We will want to solve for $N$ given a probability of 0.9 or 0.99, but there is a modification that needs to be made because we want the probability not of finding exactly 30 of theses planets, but 30 or more of them. There are two ways to calculate this probability, either by adding the probabilities of finding $n \in {[30, \infty]}$ planets, or by adding the probabilities of finding $n \in {[0, 29)}$ planets and subtracting this sum from 1. Computationally, the latter of these is the most logical, so our probability will actually be
  \begin{equation*}
    P(n \ge 30, \rho | N) = \sum_{m = 0}^{29} \left(N \choose m\right) \rho^m \left(1 - \rho\right)^{N - m},
  \end{equation*}
  and when solving we will use the substitution that $\rho = 0.1$ and $P(n \ge 30, \rho | N) = 0.9$ or $0.99$.

  Using a false position method root finding algorithm in Python, we were able to find the values of $N$ that would satisfy the above equation for our desired likelihoods. These values are
  \begin{enumerate}[a)]
  \item $N = 368$ for a 90\% likelihood
  \item $N = 435$ for a 99\% likelihood
  \end{enumerate}
